Automatically drawing vegetation maps using digital time-lapse cameras in alpine ecosystems

### 2022-06-06 paper/paper.tex
Initial commit

\usepackage{amsmath}
\setlength{\parindent}{10.5pt}
\begin{document}
\maketitle

@@

Previous studies have mainly depended on field observations, yet it is
hard to cover broad areas in alpine regions due to poor accessibility
and severe weather. Satellite, airborne, and Unmanned Aerial Vehicle
(UAV) remote sensing methods seem to be alternatives. However, satellite
imagery of alpine areas are rarely available due to cloud cover, and the
spatial resolution is not enough to observe vegetation changes at the
plant community scale. Airborne imagery can obtain high-resolution data,
but its cost becomes a bottleneck for frequent monitoring. Although UAV
methods have become a cost-effective tool for ecological monitoring
(\cite{Baena2017PLOSONE}), operating UAVs in alpine regions is
challenging due to the strong wind and harsh topology.

On the other hand, researchers have also utilized automated digital
time-lapse cameras mounted on the ground for monitoring green-leaf
phenologies in forests (\cite{Richardson2009EcolAppl}), grasslands
(\cite{Browning2017RemSen}), and alpine meadows
(\cite{IdeOguma2013EcolInfom}). Unlike satellite imagery, such cameras
provide images free of clouds and atmospheric effects. Also, they can
obtain high-resolution (i.e., sub-meter scale) and frequent (i.e., daily
or hourly) images at a meager cost. These studies set some regions of
interest (ROI) in the images and calculate the phenology index (e.g.,
excess greenness, \cite{Woebbecke1995ASAE}). Also, we can visually
interpret the vegetation distribution and its changes from such
time-lapse images. However, few studies have tackled this. This lack
seems to be because applying such time-lapse imagery in monitoring
vegetation distribution has two technical challenges.

First, unlike satellites' multispectral sensors, ordinal digital cameras
can only obtain three bands (Red, Green, and Blue), making it harder to
classify the vegetation. Second, since digital time-lapse cameras are
mounted on the ground, applying geo-spatial analysis to these images is
challenging. In other words, even if we find a vegetation change, we
cannot identify its geographic location and analyze it with topological
data. Treating ground-based time-lapse images as geographic data is
essential to utilize them in conservation planning.

This study proposes an automated method for drawing vegetation maps and
locating vegetation changes with a digital time-lapse camera by solving
these two challenges. We solved the first challenge by utilizing
temporal changes in leaf colors in vegetation classification. We also
developed an accurate method for transforming a ground-based photograph
into geographic data to tackle the second challenge. We aim to use cheap
but powerful digital time-lapse cameras in alpine ecosystem
conservation.

@@

Since ground-based photographs can provide high spatial resolution
(sub-meter scale in our dataset), we can use leaf texture as additional
information for classification. For example, we can see a glossy surface
on dwarf bamboos and a mat surface on dwarf pines. Thus, we used a small
patch (9x9 pixels square) rather than a single pixel as an input of the
model. Therefore, a model input is a time series of image patches. Our
model has two components to deal with this data structure (see Fig. 3).
First, we extracted each patch's features (e.g., texture) using
Convolutional Neural Network (CNN) layers (Fig. 4a). CNN is a neural
network specialized in recognizing spatial structures of data, such as
images. The CNN part outputs a time series of extracted features.
Second, we used Recurrent Neural Net (RNN) layers (Fig. 4b) to classify
the temporal patterns of the features extracted by the CNN part. RNN is
a neural network specialized in recognizing temporal or sequential data
dynamics, such as text and speech. Amongst many variants of RNN, we used
Long Short Time Memory (LSTM, \cite{Hochreiter1997LSTM}). Combining CNN
and RNN, our model can classify the input image-patch time series
considering the colors and textures of each patch and their temporal
patterns. This CNN-RNN architecture has also been used for audio
classification(\cite{Shi2015CRNNtext}) and scene text
recognition(\cite{Shi2015CRNNtext}). Using a Deep Learning method has
another positive side effect; mini-batch training. Since we used
high-resolution images (16 x 21M pix), the training dataset became so
immense that we could not feed them to the model at once: that consumes
the computer's RAM too much. However, using Deep Learning methods
enables us to provide them in small portions (called mini-batches) to
save RAM consumption.

@@

We set 5 classes: dwarf pine, dwarf bamboo, other vegetation, no
vegetation, and sky. Using free image annotation software (Semantic
Segmentation Editor), an expert prepared teacher data (Fig. x) for
classification.

@@

Then, we developed a novel method to convert ground-based landscape
imagery into GIS-ready geographical data. This process is called
georectification. Georectification of ground-based images has been a
difficult task, and this causes the underuse of potentially rich
information in ground-based imagery. In plain words, georectification
means aligning images into Digital Surface Models (DSMs) so that every
pixel of an image gets a geographical coordinate. We can consider a
camera as a function that transforms 3D geographical coordinates
(latitude, longitude, height) into 2D image coordinates (locations of
each pixel). So estimating the parameters of this function (such as the
camera location, pose, and the field of view), we can align an image to
a DSM. Usually, georectification has three steps: 1. Finding Ground
Control Points (GCPs) in the image. 2. Estimating camera parameters such
as camera poses and field of view using GCPs. 3. Mapping the image onto
the DSM using camera parameters.

@@

We modeled the lens distortion based on \cite{Weng1992CameraCalib} and
OpenCV's implementation (Eq. 1). Our model includes radial
(k1\textasciitilde k6), tangental (p1, p2), thin prism
(s1\textasciitilde s4) distortion, and unequal pixel aspect ratio (a1,
a2). See Weng et al.~1992 for lens distortion modeling. Also, we
recommend readers to \cite{Portenier2020Cryosphere} for camera models
without lens distortions. Thus, now we can project GCPs' geographical
coordinates into image coordinates using lens distortion parameters and
other camera parameters (the camera location, pose, and the field of
view). We optimized these (except camera location) by minimizing the
square projection error of the GCPs using the Covariance Matrix
Adaptation Evolution Strategy (CMA-ES, \cite{Hansen2003CMAES}). We could
not estimate the camera location because it makes the problem too
complicated (e.g., a telephoto taken from a distance and a wide-angle
taken from a close look similar).

\[
\begin{bmatrix}
{x}'' \\ 
{y}'' 
\end{bmatrix} 
= 
\begin{bmatrix} 
x’ \frac{1 + k_1 r^2 + k_2 r^4 + k_3 r^6}{1 + k_4 r^2 + k_5 r^4 + k_6 r^6} + 2 p_1 x’ y’ + p_2(r^2 + 2 x’^2) + s_1 r^2 + s_2 r^4 \\ 
y’ \frac{1 + a_1 + k_1 r^2 + k_2 r^4 + k_3 r^6}{1 + a_2 + k_4 r^2 + k_5 r^4 + k_6 r^6} + p_1 (r^2 + 2 y’^2) + 2 p_2 x’ y’ + s_3 r^2 + s_4 r^4 \ \end{bmatrix}
\]

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0541}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1405}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1568}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1243}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1351}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1568}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1459}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.0865}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Macro Average}, N = 5
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Weighted Average}, N = 5
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Dwarf Pine}, N = 5
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Dwarf Bamboo}, N = 5
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Other vegetation}, N = 5
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Non Vegetation}, N = 5
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{Sky}, N = 5
\end{minipage} \\
\midrule
\endhead
Precision & 0.977 (0.008) & 0.997 (0.001) & 0.977 (0.012) & 0.951
(0.013) & 0.999 (0.000) & 0.959 (0.066) & 1.000 (0.000) \\
Recall & 0.987 (0.001) & 0.997 (0.001) & 0.995 (0.003) & 0.960 (0.006) &
0.996 (0.002) & 0.984 (0.004) & 1.000 (0.000) \\
F1 Score & 0.982 (0.005) & 0.997 (0.001) & 0.986 (0.005) & 0.955 (0.004)
& 0.997 (0.001) & 0.970 (0.034) & 1.000 (0.000) \\
\bottomrule
\end{longtable}

@@

\hypertarget{the-performance-of-our-georectification-method-and-its-limitation.}{%
\subsection{The performance of our georectification method and its
limitation.}\label{the-performance-of-our-georectification-method-and-its-limitation.}}

>>>>

\usepackage{amsmath}
\setlength{\parindent}{10.5pt}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\begin{document}
\maketitle

@@

Previous studies have mainly depended on field observations, yet it is hard to cover broad areas in alpine regions due to poor accessibility and severe weather. Satellite, airborne, and Unmanned Aerial Vehicle (UAV) remote sensing methods seem to be alternatives. However, satellite imagery of alpine areas is rarely available due to cloud cover, and the spatial resolution is not enough to observe vegetation changes at the plant community scale. Airborne imagery can obtain high-resolution data, but its cost becomes a bottleneck for frequent monitoring. Although UAV methods have become a cost-effective tool for ecological monitoring (\cite{Baena2017PLOSONE}), operating UAVs in alpine regions is challenging due to the strong wind and harsh topology.

On the other hand, researchers have also utilized automated digital time-lapse cameras mounted on the ground for monitoring green-leaf phenologies in forests (\cite{Richardson2009EcolAppl}), grasslands (\cite{Browning2017RemSen}), and alpine meadows (\cite{IdeOguma2013EcolInfom}). Unlike satellite imagery, such cameras provide images free of clouds and atmospheric effects. Also, they can obtain high-resolution (i.e., sub-meter scale) and frequent (i.e., daily or hourly) images at a meager cost. These studies set some regions of interest (ROI) in the images and calculate the phenology index (e.g., excess greenness, \cite{Woebbecke1995ASAE}). Also, we can visually interpret the vegetation distribution and its changes from such time-lapse images. However, few studies have tackled this. This lack of studies seems to be because applying such time-lapse imagery in monitoring vegetation distribution has two technical challenges.

First, unlike satellites' multispectral sensors, ordinal digital cameras can only obtain three bands (Red, Green, and Blue), making it harder to classify the vegetation. Second, since digital time-lapse cameras are mounted on the ground, transforming these images into geospatial data (e.g., orthoimage) is challenging. In other words, even if we find a vegetation change, we cannot identify its geographic location and analyze it with other geographic data (e.g., topography). Treating ground-based time-lapse images as geographic data is essential to utilize them in conservation planning.

This study proposes an automated method for drawing vegetation maps and locating vegetation changes with a digital time-lapse camera by solving these two challenges. We solved the first challenge by utilizing temporal changes in leaf colors in vegetation classification. We also developed an accurate method for transforming a ground-based photograph into geographic data to tackle the second challenge. Finally, we show an example application of our method by detecting and localizing vegetation changes with an 8-years digital time-lapse imagery of the Japanese alps. We aim to use cheap but powerful digital time-lapse cameras in alpine ecosystem conservation.

@@

Since ground-based photographs can provide high spatial resolution (sub-meter scale in our dataset), we can use leaf texture as additional information for classification. For example, we can see a glossy surface on dwarf bamboos and a mat surface on dwarf pines. Thus, we used a small patch (9x9 pixels square) rather than a single pixel as an input of the model. Therefore, a model input is a time series of image patches.

Our model has two components to deal with this data structure (see Fig. 3). First, we extracted each patch's features (e.g., texture) using Convolutional Neural Network (CNN) layers (Fig. 4a). CNN is a neural network specialized in recognizing spatial structures of data, such as images. The CNN part outputs a time series of extracted features. Second, we used Recurrent Neural Net (RNN) layers (Fig. 4b) to classify the temporal patterns of the features extracted by the CNN part. RNN is a neural network specialized in recognizing temporal or sequential data dynamics, such as text and speech. Amongst many variants of RNN, we used Long Short Time Memory (LSTM, \cite{Hochreiter1997LSTM}). Combining CNN and RNN, our model can classify the input image-patch time series considering the colors and textures of each patch and their temporal patterns. This CNN-RNN architecture has also been used for audio classification(\cite{Shi2015CRNNtext}) and scene text recognition(\cite{Shi2015CRNNtext}).

Using a Deep Learning method has another positive side effect; mini-batch training. Since we used high-resolution images (16 x 21M pix), the training dataset became so immense that we could not feed them to the model at once: that consumes the computer's RAM too much. However, using Deep Learning methods enables us to provide them in small portions (called mini-batches) to save RAM consumption.

We set 5 vegetation classes: dwarf pine, dwarf bamboo, other vegetation, no vegetation, and sky. Using a free image annotation software (Semantic Segmentation Editor, Hitachi, \url{https://github.com/Hitachi-Automotive-And-Industry-Lab/semantic-segmentation-editor}), an expert drew polygons on an image of 2015 for each class. Since vegetation might be different in 2012, 2015, and 2020, applying this teacher data to the image of 2012 and 2020 may cause misclassification. However, vegetation changes are expected to occur at the edge of the plant community. Thus we shrink each polygon with several pixels before applying it to the training.

@@

@@

Then, we developed a novel method to convert ground-based landscape imagery into GIS-ready geographical data. This process is called georectification. Georectification of ground-based images has been a difficult task, and this causes the underuse of potentially rich information in ground-based imagery. In plain words, georectification means aligning images into Digital Surface Models (DSMs) so that every pixel of an image gets a geographical coordinate. We can consider a camera as a function that transforms 3D geographical coordinates (latitude, longitude, height) into 2D image coordinates (locations of each pixel). So estimating the parameters of this function (such as the camera location, pose, and the field of view), we can map an image onto a DSM. We recommend readers to \cite{Portenier2020Cryosphere} for a plain mathmatical explanation of this process. Usually, georectification has three steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Finding Ground Control Points (GCPs) in the image.\\
\item
  Estimating camera parameters such as camera poses and field of view using GCPs.\\
\item
  Mapping the image onto the DSM using camera parameters.
\end{enumerate}

@@

We modeled the lens distortion (\ref{camera_model}) based on \cite{Weng1992CameraCalib} and OpenCV's implementation, where \({x}''\) and \({y}''\) are the coordinates of each pixels after distortion and \(x'\) and \(y'\) are the coordinates before distortion. Our model includes radial (\(k1~k6\)), tangental (\(p1, p2\)), thin prism (\(s1~s4\)) distortion, and unequal pixel aspect ratio (\(a1, a2\)). See Weng et al.~1992 for the details of lens distortion modeling. Thus, now we can project GCPs' geographical coordinates into image coordinates using lens distortion parameters and other camera parameters (the camera location, pose, and the field of view). We optimized these (except camera location) by minimizing the square projection error of the GCPs using the Covariance Matrix Adaptation Evolution Strategy (CMA-ES, \cite{Hansen2003CMAES}). We could not estimate the camera location because it makes the problem too complicated (e.g., a telephoto taken from a distance and a wide-angle taken from a close look similar).

\begin{equation}
\label{camera_model}
\begin{bmatrix}
{x}'' \\ 
{y}'' 
\end{bmatrix} 
= 
\begin{bmatrix} 
x’ \frac{1 + k_1 r^2 + k_2 r^4 + k_3 r^6}{1 + k_4 r^2 + k_5 r^4 + k_6 r^6} + 2 p_1 x’ y’ + p_2(r^2 + 2 x’^2) + s_1 r^2 + s_2 r^4 \\ 
y’ \frac{1 + a_1 + k_1 r^2 + k_2 r^4 + k_3 r^6}{1 + a_2 + k_4 r^2 + k_5 r^4 + k_6 r^6} + p_1 (r^2 + 2 y’^2) + 2 p_2 x’ y’ + s_3 r^2 + s_4 r^4 \ \end{bmatrix}
\end{equation}

@@

We evaluated the performance of the vegetation classifier using a 5-fold cross-validation design. Each fold was stratified with the vegetation categories. We used three standard metrics in machine learning evaluation: precision, recall, and F1-score (\ref{metrics}), where \(TP, FP, FN\) is true positives, false positives, and false negatives, respectively.

\begin{align}
\label{metrics}
precision = \frac{TP}{TP + FP} \\
recall = \frac{TP}{TP + FN} \\
F_1 = \frac{2 \cdot precision \cdot recall}{precision + recall} \\
\end{align}

The vegetation classifier achieved a high recall and precision in all five categories \ref{tab:vegetation_cv}. Recall and precision were lowest in dwarf bamboo with the smallest training data. Even in dwarf bamboo, false positives occurred at low frequency (about 5\%), and only a tiny amount of true positives were missed by the pipeline (about 4\%). Since the weighted average of each metric was higher (more than 0.99), balancing the training dataset with vegetation categories may improve these results.

\ref{fig:vegetation} shows the products of vegetation classification. We can observe the distribution of dwarf pine and dwarf bamboo at the plant-community scale.

\begin{table}

\caption{\label{tab:unnamed-chunk-1}Cross validation result \label{tab:vegetation_cv}}
\centering
\begin{tabular}[t]{llll}
\toprule
vegetation & f1-score & precision & recall\\
\midrule
 & 0.982 & 0.977 & 0.987\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash Macro Average} & (0.005) & (0.008) & (0.001)\\
\cmidrule{1-4}
 & 0.997 & 0.997 & 0.997\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash Weighted Average} & (0.001) & (0.001) & (0.001)\\
\cmidrule{1-4}
 & 0.986 & 0.977 & 0.995\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash Dwarf Pine} & (0.005) & (0.012) & (0.003)\\
\cmidrule{1-4}
 & 0.955 & 0.951 & 0.960\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash Dwarf Bamboo} & (0.004) & (0.013) & (0.006)\\
\cmidrule{1-4}
 & 0.997 & 0.999 & 0.996\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash Other vegetation} & (0.001) & (0.000) & (0.002)\\
\cmidrule{1-4}
 & 0.970 & 0.959 & 0.984\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash Non Vegetation} & (0.034) & (0.066) & (0.004)\\
\cmidrule{1-4}
 & 1.000 & 1.000 & 1.000\\
\cmidrule{2-4}
\multirow{-2}{*}{\raggedright\arraybackslash Sky} & (0.000) & (0.000) & (0.000)\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{4}{l}{\rule{0pt}{1em}Mean (SD) of 5-Fold Cross-Varidation}\\
\end{tabular}
\end{table}

\begin{figure}
\includegraphics[width=1\linewidth]{vegetation} \caption{Vegetation classification results}\label{fig:vegetation}
\end{figure}

@@

\hypertarget{the-performance-of-the-georectification-method-and-its-limitation.}{%
\subsection{The performance of the georectification method and its limitation.}\label{the-performance-of-the-georectification-method-and-its-limitation.}}
